State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-20 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
small edits to the discussion

\subfigure[][]{\includegraphics[width=1.5in]{./Graph/ZetaConvergence.pdf}}

@@

\caption{Mean kernel estimates over 250 runs ;  true and estimated kernels are shown by solid
and dashed line respectively, the grey area shows 95\% confidence interval; Left: spatial domain; Right: frequency domain; a, b: wider basis functions; c,d: narrower basis functions.**************************Guys the bottom left figure is the power in db for skinny basis functions}
\label{fig:KernelEstimates}
\end{figure*}
\begin{figure}[th]
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the parameters ratios over 250 runs.}
\end{figure}
\label{fig:ParametersRatio}
\begin{figure}[th]
\centering
\subfigure[][]{\includegraphics[scale=0.9]{./Graph/FFTTrueField.pdf}}\label{fig:FieldFFT}
\subfigure[][]{\includegraphics[scale=0.9]{./Graph/FFTEstimatedField.pdf}}\label{fig:EstimatedFieldFFT}\\
\caption{The average FFT of the neural field; a: True field; b: Estimated field (showing the limited bandwidth).}
\label{fig:FFTTrueEstimate}
\end{figure}

  \begin{figure}
   	\begin{center}
   		\includegraphics{./Graph/MRMSE.pdf} 
   	\end{center}
   	\caption{The mean of RMSE of the estimated field over 250 runs.} 
\label{fig:RMSE}
   \end{figure}
\begin{figure*}[!th]
\centering 
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/EstimationError90.pdf}}\\
\caption{One snapshot of the spatial field; a: True field; b: Estimated field; c: The absolute error.}
\label{fig:FieldEstimate}
\end{figure*}

@@

\dean{FIRST DESCRIBE AN OVER ALL SUMMARY OF WHAT WE HAVE DONE AND WHAT IS NOVEL}
This paper presents a novel framework for model reduction to facilitate practical state and parameter estimation of neural fields. A basis function decomposition reduces the infinite dimensional neural field (continuous space), to a finite dimensional state-space model. The level of detail of the estimated field is related to the basis functions decomposition, providing a flexible framework where a trade-off can be made between physiological detail and computational complexity. By forming a state-space model, the unscented Rauch Tung Striebel Smoother was implemented for state estimation. Using the sequence of state estimates, a least squares algorithm was implemented for parameter estimation.

\dean{DISCUSS RESULTS: FREQ ANALYSIS. USE THIS TO JUSTIFY KERNEL ERRORS.}
The frequency analysis and model selection tools presented in this paper specify the sensor and field basis function arrangement required to form a state-space model that is capable of capturing the dominant cortical dynamics. These tools can be also used to better design electrophysiological recording techniques to advance model-based neural parameter estimation in order to avoid spatial aliasing. The effect of the basis function decomposition will restrict the frequency content of the estimated field. This is evident in when comparing figures~\ref{fig:FieldFFT} and~\ref{fig:EstimatedFieldFFT} where the estimated field is spatially band-limited. The spatial frequency content of the field is governed by disturbance covariance and the shape of the connectivity kernel. Therefore some information will be lost due to the field decomposition. The implication of this loss of information is the high spatial frequency content of the connectivity kernel. This is demonstrated in figure~\ref{fig:KernelEstimates} where the frequency content of the estimated connectivity kernel is less than the true kernel.

\dean{DISCUSS THAT EVEN THOUGH THE KERNEL ESTIMATE IS NOT EXACT, IT IS USEFUL TO KNOW THE RATIO OF EXCITATION TO INHIBITION AND THAT RELATIVE MEASURES CAN BE ACCURATELY TRACKED}
Even though the high frequency content of the connectivity kernel is lost through the basis function decomposition, the dominant dynamics are still captured with the estimated reduced model. This is illustrated in figure~\ref{fig:FieldEstimate}. In addition, the ratio of the kernel parameters follows a tight distribution as seen in figure~\ref{fig:ParametersRatio}. This indicates that the ratio of excitation to inhibition can be tracked with this approach.

@@

\dean{DISCUSS THE IMPLICATION OF ESTIMATING INTRACORTICAL CONNECTIVITY AND WHERE WE FIT WITHIN OTHER FUNCTIONAL CONNECTIVITY RESEARCH AREAS}
To the authors' best knowledge, this is the first paper to propose an estimation framework for intracortical connectivity from electrophysiological data. The importance of the estimation framework can be illustrated by considering that the connectivity structure at this scale can not be estimated using diffusion tensor imaging due to the near isotropic diffusion of water in grey matter~\cite{Assaf2008}. In addition, other non-model-based techniques for estimating functional connectivity, such as Granger causality~\cite{Hesse2003}, the direct transfer function~\cite{Kaminski1991} and partial directed coherence~\cite{Sameshima1999}, are restricted to the spatial scales of the recording electrodes. The model-based technique proposed in this paper estimates the connectivity structure independently of the recording electrode geometry. An advantage of the auto-regressive (AR) methods is that they allow for anisotropic connectivity, which is present in cortico-cortical fibres. It should be noted, however, that the AR and the proposed model-based methods are complementary, where connectivity is estimated at different spatial scales. Therefore, we envisage that by combining these methods a more realistic model can be acquired. A recent theoretical study~\cite{Jirsa2009} has demonstrated the importance of this multi-scale structure in generating the characteristic rhythms of the brain.

@@

\begin{itemize}
	\item What about the transient! The need for activity to perform estimation. \dean{To Learn You Must Perturb!} 
\end{itemize}

\section{Conclusion}\label{ConclusionSection}
This paper presented a novel model-based framework for patient-specific state and parameter estimation of neural fields. This framework enables a fundamental link between patient-specific physiologic data and theoretical neural field models. This link has the potential to lead to advances in our basic understanding of the dynamics of the brain at the macro/mesoscopic scale. Forming a greater understanding neural dynamics at this scale may lead to new and improved treatment options for diseases such as epilepsy.

>>>>

\subfigure[$\zeta$ convergence]{\includegraphics[width=1.5in]{./Graph/ZetaConvergence.pdf}}

@@

\caption{Results for connectivity estimate. In all the subplots the true kernel is show by the solid line, the mean kernel estimate (over 250 realisations) is shown by the dashed line and the 95~\% confidence interval is shown by the shaded grey region. \textbf{a.} Comparison of kernel estimates to true kernel in the spatial domain. \textbf{b.} Comparison of the true kernel to the estimated kernel in the frequency domain. \dean{Note: needs to be changed to absolute values if we include it.} \textbf{c.} Same as part a., except narrower wider (in space) basis functions where used in the model reduction. \textbf{d.} Same as part b., except narrower wider (in space) basis functions where used in the model reduction. \textbf{e.} Frequency domain (power dB) representation of the true kernel and the estimated kernel (using the narrower basis functions). \dean{Note: Power and dB needs to be non-italicised.}}
\label{fig:KernelEstimates}
\end{figure*}
\begin{figure}[th]
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 250 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20).}
\end{figure}
\label{fig:ParametersRatio}
\begin{figure}[th]
\centering
\subfigure[][]{\label{fig:FieldFFT}\includegraphics[scale=0.9]{./Graph/FFTTrueField.pdf}}
\subfigure[][]{\label{fig:EstimatedFieldFFT}\includegraphics[scale=0.9]{./Graph/FFTEstimatedField.pdf}}\\
\caption{Spatial frequency of the neural fields. \textbf{a.} The average (over time) spatial frequency of the neural field. \textbf{b.} The average (over time) spatial frequency of the estimated neural field resulting from the basis function decomposition. As a result of the band limiting properties of the sensors and basis functions the reconstructed shows a narrower bandwidth.}
\label{fig:FFTTrueEstimate}
\end{figure}

  \begin{figure}
   	\begin{center}
   		\includegraphics{./Graph/MRMSE.pdf} 
   	\end{center}
   	\caption{The mean of average RMSE of the estimated field over 250 realisations (solid line) and the 95~\% confidence interval (grey region). \dean{David suggested for us to use percentage error, and use a hard threshold of points close to zero to stop problems with divide by zeros. Also, it might be good to show the data points along the line, so we can see the number of iterations it takes to stabilise.}} 
\label{fig:RMSE}
   \end{figure}
\begin{figure*}[!th]
\centering 
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/EstimationError90.pdf}}\\
\caption{Examples of the neural fields. \textbf{a.} The true field from the non-reduced model the was used to generate data. \textbf{b.} The reconstructed estimated field of the reduced model. \textbf{c.} The absolute error between the true and reconstructed fields.}
\label{fig:FieldEstimate}
\end{figure*}

@@

\dean{FIRST DESCRIBE AN OVER ALL SUMMARY OF WHAT WE HAVE DONE AND WHAT IS NOVEL}
This paper presents a novel framework for model reduction to facilitate practical state and parameter estimation of spatio-temporal neural fields. A basis function decomposition reduces the infinite dimensional neural field (continuous space), to a finite dimensional state-space model. The level of detail of the estimated field is related to the basis functions decomposition, providing a flexible framework where a trade-off can be made between physiological detail and computational complexity. Using the state-space model, the unscented Rauch Tung Striebel Smoother was applied for state estimation. Using the sequence of state estimates, a least squares algorithm was implemented for parameter estimation.

\dean{DISCUSS RESULTS: FREQ ANALYSIS. USE THIS TO JUSTIFY KERNEL ERRORS.}
The frequency analysis and model selection tools presented in this paper specify the sensor and field basis function arrangement required to form a state-space model that is capable of capturing the dominant cortical dynamics. In addition to providing conditions to ensure the reduced model can capture the significant spectral properties of the field, these tools can be used to better design electrophysiological recording techniques avoid spatial aliasing. The effect of the basis function decomposition will restrict the frequency content of the estimated field. This is evident in when comparing figures~\ref{fig:FieldFFT} and~\ref{fig:EstimatedFieldFFT} where the estimated field is spatially band-limited. The spatial frequency content of the field is governed by disturbance covariance and the shape of the connectivity kernel. Therefore some information will be lost due to the field decomposition. The implication of this loss of information is the high spatial frequency content of the connectivity kernel. This is demonstrated in figure~\ref{fig:KernelEstimates} where the frequency content of the estimated connectivity kernel is less than the true kernel.

\dean{DISCUSS THAT EVEN THOUGH THE KERNEL ESTIMATE IS NOT EXACT, IT IS USEFUL TO KNOW THE RATIO OF EXCITATION TO INHIBITION AND THAT RELATIVE MEASURES CAN BE ACCURATELY TRACKED}
Even though the some high spatial frequency content of the connectivity kernel is lost through the basis function decomposition, the dominant dynamics are still captured with the estimated reduced model. This is illustrated in figure~\ref{fig:FieldEstimate}. In addition, the ratio of the kernel parameters follows a tight distribution as seen in figure~\ref{fig:ParametersRatio}. This indicates that the ratio of excitation to inhibition can be tracked with this approach.

@@

\dean{DISCUSS THE IMPLICATION OF ESTIMATING INTRACORTICAL CONNECTIVITY AND WHERE WE FIT WITHIN OTHER FUNCTIONAL CONNECTIVITY RESEARCH AREAS}
To the authors' best knowledge, this is the first paper to propose an estimation framework for intracortical connectivity from electrophysiological data. The importance of the estimation framework can be illustrated by considering that the connectivity structure at this scale can not be estimated using diffusion tensor imaging due to the near isotropic diffusion of water in grey matter~\cite{Assaf2008}. In addition, other non-model-based techniques for estimating functional connectivity, such as Granger causality~\cite{Hesse2003}, the direct transfer function~\cite{Kaminski1991} and partial directed coherence~\cite{Sameshima1999}, are restricted to the spatial scales of the recording electrodes. The model-based technique proposed in this paper estimates the connectivity structure of the number of sensors. An advantage of the auto-regressive (AR) methods is that they allow for anisotropic connectivity, which is present in cortico-cortical fibres. It should be noted, however, that the AR and the proposed model-based methods are complementary, where connectivity is estimated at different spatial scales. Therefore, we envisage that by combining these methods a more realistic model can be acquired. A recent theoretical study~\cite{Jirsa2009} has demonstrated the importance of this multi-scale structure in generating the characteristic rhythms of the brain.

@@

\begin{itemize}
	\item What about the transient! The need for activity to perform estimation. \dean{To Learn You Must Perturb!} 
\end{itemize}
% 
% \section{Conclusion}\label{ConclusionSection}
% This paper presented a novel model-based framework for patient-specific state and parameter estimation of neural fields. This framework enables a fundamental link between patient-specific physiologic data and theoretical neural field models. This link has the potential to lead to advances in our basic understanding of the dynamics of the brain at the macro/mesoscopic scale. Forming a greater understanding neural dynamics at this scale may lead to new and improved treatment options for diseases such as epilepsy.
